Learning Discrete Bayesian Networks from Continuous Data

### 2015-12-02 BN_discretizer_paper.tex
misc

Table~\ref{table:auto_disc_table_1} lists the discretization edges and mean log-likelihoods under \num{10}-fold cross validation of the discrete Bayesian network resulting from the two discretization methods.
The MDL method does not produce any discretization edges, assigning one continuous interval to each continuous variable, and produces the result with the lower likelihood.
The reason why MDL produces fewer discretization edges is discussed in Section~\ref{subsec:discuss_exp}.
Two examples of Bayesian networks where MDL produces comparable results to the Bayesian method are given in the Appendix.

@@

Two examples of Bayesian networks where MDL produces comparable results to the Bayesian method are given in this section. The first example is on the Auto MPG dataset. Instead of testing on the network obtained by prediscretization and structure learning, the tested network is obtained by Algorithm \ref{alg:structure_learn}, which can be considered better-structured than the network in Figure \ref{fig:auto_graph_1}. The second example is on the IRIS dataset \citep{Lichman_2013}. In the work of \cite{Friedman_1996}, the IRIS dataset was tested on the navie Bayes and the prediction accuracy was shown. The same test is reproduced to compare the Bayesian method with the MDL discretization method.

\subsection{Example 1}
The Bayesian network structure in Figure~\ref{fig:auto_graph_3} and the discretization results in Table~\ref{table:auto_disc_table_3} demonstrate that the MDL method does not always produce fewer discretization intervals than the Bayesian approach and that the MDL method can produce discretization policies with comparable likelihood.
The Bayesian network in Figure~\ref{fig:auto_graph_3} was obtained by running Algorithm~\ref{alg:structure_learn} fifty times on the Auto MPG dataset with a maximum of two parents per variable.
The network is better-structured than the one in Figure~\ref{fig:auto_graph_1} because the discretization policies were simultaneously learned with the network structure.
Many edges appear in both discretization results and are bolded in Table~\ref{table:auto_disc_table_3}.

@@

\begin{table}[ht]
  \centering
  \caption{
    The discretization policies for the Auto MPG dataset with the fixed structure shown in Figure~\ref{fig:auto_graph_3}.
    % The first five rows list the discretization edges for each continuous variable.
    % The last row shows the mean cross-validated log-likelihood under each policy.
    The methods have the same discretization for variable $A$, and the log-likelihood for the Bayesian method is only slightly better than that for the MDL method.
  }
\input{Table4.tex}
  \label{table:auto_disc_table_3}
\end{table}

\subsection{Example 2}
The Iris dataset contains variables related to the morphologic variation of three Iris flower species.
The dataset has \num{150} samples over five variables.
Only variables $E$ is discrete-valued, which corresponds to the three species.
The following experiment was conducted on the naive Bayes: variable $E$ is the parent of variable $A$, $B$, $C$, and $D$. Table \ref{table:iris_table_1} shows the discretization edges and the prediction accuracy of the learned classifiers using 10-fold cross validation. The MDL discretization method and the Bayesian method have same discretization result: all discretization edges coincide. In addition, both methods have $94 \% $ accuarcy, which matches the number in \cite{Friedman_1996}.

\begin{table}[ht]
  \centering
  \caption{
    The discretization policies for the Iris dataset with the naive Bayes structure. 
    Two methods produce same discretization policy on each continuous variable and have same prediction accuracy.
  }
\input{Table5.tex}
  \label{table:iris_table_1}
\end{table}
\end{document}
% end of file template.tex

>>>>

Table~\ref{table:auto_disc_table_1} lists the discretization edges and mean log-likelihoods under \num{10}-fold cross validation of the discrete Bayesian network resulting from the two discretization methods.
The MDL method does not produce any discretization edges, assigning one continuous interval to each continuous variable, and produces the result with the lower likelihood.
The reason why MDL produces fewer discretization edges is discussed in Section~\ref{subsec:discuss_exp}.
Two examples of MDL producing comparble results to the Bayesian method are given in the Appendix.

@@

The Bayesian network structure in Figure~\ref{fig:auto_graph_3} and the discretization results in Table~\ref{table:auto_disc_table_3} demonstrate that the MDL method does not always produce fewer discretization intervals than the Bayesian approach and that the MDL method can produce discretization policies with comparable likelihood.
The Bayesian network in Figure~\ref{fig:auto_graph_3} was obtained by running Algorithm~\ref{alg:structure_learn} fifty times on the Auto MPG dataset with a maximum of two parents per variable.
The network is structured more favorably with regards to discretization than the network in Figure~\ref{fig:auto_graph_1}, as the discretization policies were simultaneously learned with the network structure, resulting in higher data likelihood.
Many edges appear in both discretization results and are bolded in Table~\ref{table:auto_disc_table_3}.

@@

\begin{table}[ht]
  \centering
  \caption{
    The discretization policies for the Auto MPG dataset with the fixed structure shown in Figure~\ref{fig:auto_graph_3}.
    % The first five rows list the discretization edges for each continuous variable.
    % The last row shows the mean cross-validated log-likelihood under each policy.
    The methods have the same discretization for variable $A$, and the log-likelihood for the Bayesian method is only slightly better than that for the MDL method.
  }
\input{Table4.tex}
  \label{table:auto_disc_table_3}
\end{table}
\end{document}
% end of file template.tex
